### 2015-06-05 Manuscript/Missing_mammals-draft.tex
Edits to data collection

% NC: If there are word count problems we can ditch the last two sentences.
%---------------------------------------------
%
%       METHODS
%
%---------------------------------------------

\section{Materials and Methods}
\subsection{Data collection and standardisation}
We downloaded all cladistic matrices containing any living or/and fossil mammal taxa from three major public databases: Morphobank (\texttt{http://www.morphobank.org/}) \citep{morphobank}, Graeme Lloyd's website (\texttt{graemetlloyd.com/matrmamm.html}) and Ross Mounce's GitHub repository (\texttt{https://github.com/rossmounce/cladistic-data}).
We also performed a systematic Google Scholar search for matrices that were not uploaded to these databases (see Supplementary Materials for a detailed description of the search procedure).
In total, we downloaded @256 % Number of matrices
matrices containing unique @9411 % Number of unique OTUs
operational taxonomic units (OTUs).
In this study, we refer to OTUs rather than species since the entries in the downloaded matrices were not standardized and ranged from specific specimen names to family level.
When possible, we considered the OTUs at their lowest valid taxonomic level (species) but some OTUs where only at a higher valid taxonomic level (e.g. genus or family).
Therefore for some orders, we sampled more genus than species (Table \ref{Table_morpho_taxa_proportion}).

To select the lower valid taxonomic level, we standardised the taxonomy of our OTUs by fixing invalid binomial inputs to match with standard taxonomic nomenclature (e.g., \textit{H. sapiens} was transformed to \textit{Homo sapiens}).
We designated as ``living'' all the OTUs that where either present in the phylogeny of Fritz \textit{et al}. \citep{FritzTree} or the taxonomy of Wilson \& Reeder \citep{wilson2005mammal}, and designated as ``fossil'' all the OTUs that were present in the Paleobiology database (\texttt{https://paleobiodb.org/}).
For OTUs that did not appear in these three sources we first decomposed the name (i.e. \textit{Homo$\_$sapiens} became \textit{Homo} and \textit{sapiens}) and tried to match the first element with a higher taxonomic level (family, genus etc.).
Any OTUs that still had no matches in the sources above were designated as non-applicable (NA; see Supplementary Material for more details).

The number of character per matrices ranged from @6 to @4541.
This range led to difficulties in comparing the available data between matrices because (1) small matrices have less chance of having character overlapping with other matrices \citep{wagner2000} and (2) small matrices are more likely to contain a higher proportion of specific characters non-applicable across large clades \citep{Brazeau2011} (e.g. ``antler ramifications'' is a character that is only applicable to Cervidae not all mammals).
Therefore we selected only matrices containing at least 100 cladistic characters per OTU.
This arbitrary threshold was chosen to correspond with the number of characters used in \citep{GuillermeCooper} and \citep{harrisonamong-character2014}.

We extracted @1422 unique living mammal OTUs from the @256 matrices with a minimum of 6 and a maximum of 4541 coded cladistic characters per OTU.
After removing all the matrices with less than 100 cladistic characters, the number of extracted living mammals OTUs was down to @815 from @number matrices.

>>>>

% NC: If there are word count problems we can ditch the last two sentences.
%---------------------------------------------
%
%       METHODS
%
%---------------------------------------------
\section{Materials and Methods}
\subsection{Data collection and standardisation}
We downloaded all cladistic matrices containing any living and/or fossil mammal taxa from three major public databases (accessed @th of July 2015): Morphobank \citep[\texttt{http://www.morphobank.org/};][]{morphobank}, Graeme Lloyd's website (\texttt{graemetlloyd.com/matrmamm.html}) and Ross Mounce's GitHub repository (\texttt{https://github.com/rossmounce/cladistic-data}).
We also performed a systematic Google Scholar search (accessed @th of July 2015) for matrices that were not uploaded to these databases (see Supplementary Materials for a detailed description of the search procedure).
In total, we downloaded @256 % Number of matrices
matrices containing @9411 unique % Number of unique OTUs
operational taxonomic units (OTUs).
In this study, we refer to OTUs rather than species since the entries in the downloaded matrices were not standardized and ranged from specific specimen names %NC: Do you mean names of individual specimens or subspecies?
to the family-level.
Where possible, we considered OTUs at their lowest valid taxonomic level (i.e. species) but some OTUs were only valid at a higher taxonomic level (e.g. genus or family).
Therefore for some orders, we sampled more genera than species (Table \ref{Table_morpho_taxa_proportion}).

To select the lowest valid taxonomic level for each OTU, we standardised their taxonomy by correcting species names so they matched standard taxonomic nomenclature (e.g., \textit{H. sapiens} was transformed to \textit{Homo sapiens}).
We designated as ``living'' all OTUs that were either present in the phylogeny of \citet{FritzTree} or the taxonomy of \citet{wilson2005mammal}, and designated as ``fossil'' all OTUs that were present in the Paleobiology database (\texttt{https://paleobiodb.org/}).
For OTUs that did not appear in these three sources, we first decomposed the name (i.e. \textit{Homo sapiens} became \textit{Homo} and \textit{sapiens}) and tried to match the first element with a higher taxonomic level (family, genus etc.).
Any OTUs that still had no matches in the sources above were designated as non-applicable (NA; see Supplementary Material for more details).

The number of characters in each matrix ranged from @6 to @4541.
Matrices with few characters % NC: matrices can be small in two ways - with few characters and with few species. Here you just mention few characters so I've made it more specific.
are problematic when comparing available data among matrices because (1) they have less chance of having characters that overlap with those of other matrices \citep{wagner2000} and (2) they are more likely to contain a higher proportion of specific characters that are not-applicable across large clades \citep[][e.g. ``antler ramifications'' is a character that is only applicable to Cervidae not all mammals]{Brazeau2011}.
% NC: Is this true really? Does it always follow that if you've very few characters you also have very specific characters? Maybe so...
Therefore we selected only matrices containing \textgreater 100 characters per OTU.
This threshold was chosen to correspond with the number of characters used in \citet{GuillermeCooper} and \citet{harrisonamong-character2014}.

After removing all matrices with \textless 100 characters, we retained @1422 unique living mammal OTUs from @256 matrices for our analyses.
